EuPPollNet: A European database of plant-pollinator networks

### 2024-08-20 Manuscript/EuPPollNet_Ms.qmd
edits

The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent. As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; https://doi.org/10.3030/101003476), first, data was directly requested from members of the Safeguard project. Second, the request was extended to data owners outside of the project. These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching for studies on Google Scholar of under-represented regions within the database. To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting them as the contact between a pollinator and the reproductive structure of a specific plant being sampled (i.e., phyto-centric networks).

@@

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure 1}. (\\textbf{a}) Locations of the studies in EuPPollNet showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study. The sizes of these shapes are proportional to the respective species counts. For visualization purposes, we have focused only on the European region with studies and selected a single location per study. (\\textbf{b}) Number of studies by year in EuPPollNet. (\\textbf{c}) Number of species and interactions across the four main pollinator orders in EuPPollNet."}

@@

#Plot
#p1 = ggplot(poll_spread, 
#aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
#geom_bar(stat = "identity", 
#           color = "#D55E00",
#           lwd = 0.015, fill = "#D55E00") +
#ylab("Shared pollinators (%)") +
#theme_minimal()+
#theme(axis.text.x = element_blank(),
#  axis.ticks.x=element_blank(),
#  axis.ticks.y = element_line(size = 0.2),
#  plot.title = element_text(size = 8, face = "bold"), 
#  axis.text.y = element_text(size=5),
#  panel.grid.major.x = element_blank(),
#  panel.grid.minor.y = element_blank(),
#  panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
#xlab(NULL) +
#ylab(NULL) + 
#ggtitle("Pollinators")+
#geom_vline(xintercept = quantile_50, linetype="dashed", 
#color = "black", size=0.25) + 
#annotate(geom="text", x=1285, y=0.83, label="50th percentile",          color="black", size = #1.5) +
#annotate(geom="text", x=20, y=0.88, label=most_common_poll[1],         color="black",  #fontface = 'italic', size = 1.5)+
#scale_y_continuous(breaks = seq(0, 0.75, by= 0.25), expand = c(0, 0)) + 
#annotate(geom="text", x=-500, y=1.15, label="(c)",color="black", size = 3.5) +
#coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),ylim=c(0, #max(poll_spread$Percent_total)), clip="off")
#
##----------------#
##Plants
##----------------#
#plant_spread = data %>% 
#select(Plant_rank, Plant_accepted_name, Study_id) %>% 
#group_by(Study_id) %>% 
#filter(Plant_rank == "SPECIES") %>% 
#distinct(Plant_accepted_name) %>% 
#ungroup() %>% 
#count(Plant_accepted_name) %>% 
#rename(n_studies_shared = n) %>% 
#arrange(-n_studies_shared) %>% 
#mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
#  
##Seelect quantile 50% from the total number of species
#spp_number = plant_spread %>%  
#select(Plant_accepted_name) %>% 
#n_distinct()
#df = seq(1, spp_number)
#quantile_50 = quantile(df, 0.5)
#
##Obtain most common plant for plotting
#most_common_plant = plant_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
#select(Plant_accepted_name) %>% pull() 
#
#Plot
#p2 = ggplot(plant_spread, 
#aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
#geom_bar(stat = "identity", 
#           color = "#009E73",
#           lwd = 0.015, fill = "#009E73") +
#theme_minimal()+
#theme(axis.text.x = element_blank(),
#         axis.ticks.x=element_blank(),
#         axis.ticks.y = element_line(size = 0.2),
#         axis.title.x = element_text(face = "bold", size = 7),            plot.title = #element_text(size = 8, face = "bold"),
#         axis.text.y = element_text(size=5),
#         panel.grid.major.x = element_blank(),
#         panel.grid.minor.y = element_blank(),
#         panel.grid.major.y = element_line(size=0.2, linetype =  2),
#      axis.title.y = element_text(face= "bold", size= 7, hjust= -0.01, vjust = 1)) +
#xlab("Species") +
#ylab("Occurrence across studies") + 
#ggtitle("Plants") + 
#geom_vline(xintercept = quantile_50, linetype="dashed", 
#          color = "black", size=0.25) + 
#annotate(geom="text", x=835, y=0.7, label="50th percentile",color="black", size = 1.5) + 
#annotate(geom="text", x=20, y=0.75, label=most_common_plant[1]
#      ,color="black",  fontface = 'italic', size = 1.5)+
#annotate(geom="text", x=20, y=-0.08, label="Common spp."
#      ,color="black",  size = 1.5) +
#annotate(geom="text", x=1200, y=-0.08, label="\'Rare\' spp."
#      ,color="black",  size = 1.5) + 
#scale_y_continuous(breaks = seq(0, 0.6, by= 0.2), expand = c(0, 0), labels = #label_number(accuracy = 0.01)) +
#coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),ylim=c(0, #max(plant_spread$Percent_total)), clip="off")
#

@@

main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
#Select main orders
data1 = data %>% 
select(Pollinator_order, Pollinator_accepted_name) %>% 
filter(Pollinator_order %in% main_orders) %>% 
group_by(Pollinator_order) %>% 
summarise(Species = n_distinct(Pollinator_accepted_name)/n_distinct(.),
          Interactions = length(Pollinator_accepted_name)/ nrow(.))

# Combine Interactions and Species into a single column
data_long = pivot_longer(data1, cols = c(Interactions, Species), names_to = "Variable", values_to = "Value")

@@

#Plot
p3 = ggplot(data_long, aes(x = Variable, y = Value, fill = Pollinator_order)) + 
geom_bar(stat = "identity", width = 0.7) +
theme_classic() +
scale_fill_viridis(discrete=TRUE, direction = -1, name="Pollinator order") +
xlab(NULL) +
ylab("Percentage") + 
scale_y_continuous(labels = percent_format(), expand = c(0, 0)) +
theme(axis.text = element_text(size=6), 
     axis.title.y = element_text(face = "bold", size = 7, vjust = -2),  
     axis.title.x = element_text(face="bold", size = 7),
     legend.position = "top",
     legend.title = element_text(size = 5, face = "bold"),
     legend.text = element_text(size = 5),
     legend.key.size = unit(3, "mm"),
     legend.margin=margin(0,0,0,0),
     legend.box.margin=margin(-8,-5,-5,-5)) +
guides(fill=guide_legend(nrow=2,byrow=TRUE))+
ggtitle("(c)") +
theme(plot.title = element_text(size = 10,hjust = -0.3, vjust = -6))

@@

The completeness of the EuPPollNet database was evaluated by exploring the rarefied accumulation curves of plant and pollinator species and their interactions across the different networks. In addition, an outstanding question in ecology is how many pollinators are required to pollinate flowering plants [e.g., @kleijn2015 for crops]. To that end, we also calculated the accumulation curve of pollinator species with an increasing number of plant species. The rarefied and extrapolated sampling curves were obtained using the **iNEXT** package [@hsieh2016]. The different rarefied curves were complemented with 100 boostrapped accumulation curves.

>>>>

The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent as defined by the European Environment Agency [@stanners1995]. As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; https://doi.org/10.3030/101003476), first, data was directly requested from members of the Safeguard project. Second, the request was extended to data owners outside of the project. These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching for studies on Google Scholar of under-represented regions within the database. To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting them as the contact between a pollinator and the reproductive structure of a specific plant being sampled (i.e., phyto-centric networks).

@@

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure 1}. (\\textbf{a}) Locations of the studies in EuPPollNet across the European continent, showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study. The sizes of these shapes are proportional to the respective species counts. For visualization purposes, we have selected only a single location per study. (\\textbf{b}) Number of studies by year in EuPPollNet. (\\textbf{c}) Proportion of species and interactions across the four main pollinator orders in EuPPollNet, excluding interactions from \\textit{Apis mellifera}. The total number of species and interactions is indicated in parentheses at the bottom."}

@@

@@

main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")

#Calculate spp number and interactions separately
data1_spp = data %>% 
select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
filter(Pollinator_order %in% main_orders) %>% 
filter(Pollinator_rank == "SPECIES") %>% 
group_by(Pollinator_order) %>% 
summarise(Species = n_distinct(Pollinator_accepted_name)/n_distinct(.))

data1_int = data %>% 
filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
filter(Pollinator_order %in% main_orders) %>% 
#filter(Pollinator_rank == "SPECIES") %>% 
group_by(Pollinator_order) %>% 
summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.))

data1 = left_join(data1_spp, data1_int)

#Select main orders
spp = data %>% 
select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
filter(Pollinator_order %in% main_orders) %>% 
filter(Pollinator_rank == "SPECIES") %>% 
summarise(Species = n_distinct(Pollinator_accepted_name)) %>% 
pull()
#Select main orders
int = data %>% 
filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
filter(Pollinator_order %in% main_orders) %>% 
#filter(Pollinator_rank == "SPECIES") %>% 
summarise(Interactions = length(Pollinator_accepted_name)) %>% 
pull()

#Combine Interactions and Species into a single column
data_long = pivot_longer(data1, cols = c(Interactions, Species), names_to = "Variable", values_to = "Value")

@@

#Plot
p3 = ggplot(data_long, aes(x = Variable, y = Value, fill = Pollinator_order)) + 
geom_bar(stat = "identity", width = 0.7) +
theme_classic() +
scale_fill_viridis(discrete=TRUE, direction = -1, name="Pollinator order") +
xlab(NULL) +
ylab("Percentage") + 
scale_y_continuous(labels = percent_format(), expand = c(0, 0)) +
theme(axis.text = element_text(size=6), 
     axis.title.y = element_text(face = "bold", size = 7, vjust = -2),  
     axis.title.x = element_text(face="bold", size = 7),
     legend.position = "top",
     legend.title = element_text(size = 5, face = "bold"),
     legend.text = element_text(size = 5),
     legend.key.size = unit(3, "mm"),
     legend.margin=margin(0,0,0,0),
     legend.box.margin=margin(-8,-5,-5,-5)) +
guides(fill=guide_legend(nrow=2,byrow=TRUE))+
ggtitle("(c)") +
theme(plot.title = element_text(size = 10,hjust = -0.3, vjust = -6)) +
scale_x_discrete(labels = c("Species\n (2172)", "Interactions\n (358,049)"))

@@

The completeness of the EuPPollNet database was evaluated by exploring the rarefied accumulation curves of plant and pollinator species and their interactions across the different networks. In addition, we computed the  accumulation curve of pollinator species with an increasing number of plant species as an indicator for how many pollinator species are likely responsible for the pollination of flowering plants [e.g., @kleijn2015 for crops]. The rarefied and extrapolated sampling curves were obtained using the **iNEXT** package [@hsieh2016]. The different rarefied curves were complemented with 100 bootstrapped accumulation curves.

### 2024-08-20 Manuscript/Figures_and_Tables/1_Figure1.qmd
edits

#Load data
data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")

@@

main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
#Select main orders
data1 = data %>% 
select(Pollinator_order, Pollinator_accepted_name) %>% 
filter(Pollinator_order %in% main_orders) %>% 
group_by(Pollinator_order) %>% 
summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.),
          Species = n_distinct(Pollinator_accepted_name)/n_distinct(.))

# Combine Interactions and Species into a single column
data_long = pivot_longer(data1, cols = c(Interactions, Species), names_to = "Variable", values_to = "Value")

>>>>

#Load data
data = readRDS("../../Data/3_Final_data/Interaction_data.rds")

@@

main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")

#Calculate spp number and interactions separately
data1_spp = data %>% 
select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
filter(Pollinator_order %in% main_orders) %>% 
filter(Pollinator_rank == "SPECIES") %>% 
group_by(Pollinator_order) %>% 
summarise(Species = n_distinct(Pollinator_accepted_name)/n_distinct(.))

data1_int = data %>% 
filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
filter(Pollinator_order %in% main_orders) %>% 
#filter(Pollinator_rank == "SPECIES") %>% 
group_by(Pollinator_order) %>% 
summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.))

data1 = left_join(data1_spp, data1_int)

#Select main orders
spp = data %>% 
select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
filter(Pollinator_order %in% main_orders) %>% 
filter(Pollinator_rank == "SPECIES") %>% 
summarise(Species = n_distinct(Pollinator_accepted_name)) %>% 
pull()
#Select main orders
int = data %>% 
filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
filter(Pollinator_order %in% main_orders) %>% 
#filter(Pollinator_rank == "SPECIES") %>% 
summarise(Interactions = length(Pollinator_accepted_name)) %>% 
pull()

#Combine Interactions and Species into a single column
data_long = pivot_longer(data1, cols = c(Interactions, Species), names_to = "Variable", values_to = "Value")
